---
output: 
  md_document:
    variant: markdown_github
---

[![Build Status](https://travis-ci.org/rmcd1024/derivmkts.svg?branch=master)](https://travis-ci.org/rmcd1024/derivmkts)[![](http://www.r-pkg.org/badges/version/derivmkts)](http://www.r-pkg.org/pkg/derivmkts)


```{r, echo=FALSE}
library(knitr)
opts_chunk$set(comment=NA)
library(derivmkts)
```

## Introduction

This is a collection of option pricing functions for a course in
financial derivatives. The names of the functions mostly match those
in my book *Derivatives Markets*, which explains the package name.
This is the initial version; I expect to add and refine functions over
time. Depending upon the level of the course, this package may or may
not be helpful for students. I hope that at a minimum it will be
helpful for instructors.

There are of course other option pricing packages in R, notably
RQuantLib and fOptions. I don't claim to add significant pricing
functionality to those. This package does, however, have a few aspects
that might be unique, which I describe below.

The package includes functions for computing

* Black-Scholes prices and greeks for European options

* Option pricing and plotting using the binomial model

* Barrier options

* Pricing of options with jumps using the Merton model

* Analytical and Monte Carlo pricing of Asian options 

## Things of note

### Calculation of Greeks

I have tried to make calculation of Greeks easy. The
function `greeks()` accepts an option pricing function call as an
argument, and returns a vectorized set of greeks for any pricing
function that uses standard input names (i.e., the asset price is `s`,
the volatility is `v`, etc.). 

As an example, the following two calculations will produce identical
output, computing the full complement of greeks for a call, for three
strike prices. You can access the delta values as, for example,
`x['Delta', ]`.


```{r}
x1 <- greeks(bscall(s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0))
x1
```

<!--
As alternatives that may become deprecated, `bsopt()` simultaneously
computes prices and Greeks for European calls and puts and
`greeks2()` uses a different calling convention than `greeks()`

The function
```{r, eval=FALSE}
x2 <- greeks2(bscall, s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0)
y <- bsopt(s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0)
```

will do the same for both calls and puts simultanteously. The delta
values for the call would be `y[['Call']]['Delta', ]`
-->

My favorite example is this:

```{r}
k <- 100; r <- 0.08; v <- 0.30; tt <- 2; d <- 0
S <- seq(.5, 250, by=.5)
y <- list(Call=greeks(bscall(S, k, v, r, tt, d)),
          Put=greeks(bsput(S, k, v, r, tt, d))
          )
par(mfrow=c(4, 4))  ## create a 4x4 plot
par(mar=c(2,2,2,2)) ## shrink some of the margins
for (i in names(y)) {
    for (j in rownames(y[[i]])) {  ## loop over greeks
        plot(S, y[[i]][j, ], main=paste(i, j), ylab=j, type='l')
    }
}
```

This small bit of code computes and plots all call and put Greeks for
500 options. This is 16 plots in all.  It's a great illustration of
how powerful R can be.

### Binomial calculations

#### binomopt

By default the binomopt function returns the price of an American
call. In adddition, with:

* `putopt=TRUE` it returns the price of an American
put.

*  `returngreeks=TRUE` it returns a subset of the Greeks
along with the binomial parameters.

* `returntrees=TRUE` it returns as a list all of the above plus the full binomial tree (\$stree), the probability of reaching each node (\$probtree), whether or not the option is exercised at each node  (\$exertree), and the replicating portfolio at each node ($deltatree and $bondtree).

Here is an example illustrating everything that the function can return:
```{r, echo=FALSE, eval=FALSE}
x <- binomopt(41, 40, .3, .08, 1, 0, 3, putopt=TRUE, returntrees=TRUE)
x
$price
   price 
3.292948 

$greeks
       delta        gamma        theta 
-0.331656818  0.037840906 -0.005106465 

$params
         s          k          v          r         tt          d      nstep 
41.0000000 40.0000000  0.3000000  0.0800000  1.0000000  0.0000000  3.0000000 
         p         up         dn          h 
 0.4568067  1.2212461  0.8636926  0.3333333 

$oppricetree
         [,1]      [,2]     [,3]      [,4]
[1,] 3.292948 0.7409412 0.000000  0.000000
[2,] 0.000000 5.6029294 1.400911  0.000000
[3,] 0.000000 0.0000000 9.415442  2.648727
[4,] 0.000000 0.0000000 0.000000 13.584345

$stree
     [,1]     [,2]     [,3]     [,4]
[1,]   41 50.07109 61.14913 74.67813
[2,]    0 35.41139 43.24603 52.81404
[3,]    0  0.00000 30.58456 37.35127
[4,]    0  0.00000  0.00000 26.41565

$probtree
     [,1]      [,2]      [,3]       [,4]
[1,]    1 0.4568067 0.2086723 0.09532291
[2,]    0 0.5431933 0.4962687 0.34004825
[3,]    0 0.0000000 0.2950590 0.40435476
[4,]    0 0.0000000 0.0000000 0.16027409

$exertree
      [,1]  [,2]  [,3]  [,4]
[1,] FALSE FALSE FALSE FALSE
[2,] FALSE FALSE FALSE FALSE
[3,] FALSE FALSE  TRUE  TRUE
[4,] FALSE FALSE FALSE  TRUE

$deltatree
           [,1]        [,2]       [,3]
[1,] -0.3316568 -0.07824964  0.0000000
[2,]  0.0000000 -0.63298582 -0.1712971
[3,]  0.0000000  0.00000000 -1.0000000

$bondtree
         [,1]      [,2]      [,3]
[1,] 16.89088  4.658986  0.000000
[2,]  0.00000 28.017840  8.808828
[3,]  0.00000  0.000000 38.947430
```

#### binomplot

This function plots the binomial tree, providing a visual depiction of
the nodes, the probability of reaching each node, and whether exercise
occurs at that node.

### Galton board or quincunx

The [Galton board](http://mathworld.wolfram.com/GaltonBoard.html) is a
pegboard that illustrates the central limit theorem. Balls drop from
the top and randomly fall right or left, providing a physical
simulation of a binomial distribution. (My physicist brother-in-law
tells me that real-life Galton boards don't typically generate a
normal distribution because, among other things, balls acquire
momentum in the direction of their original travel. The distribution
is thus likely to be fatter-tailed than normal.)

You can see the Galton board in action with `quincunx()`.

## Feedback

Please feel free to contact me with bug reports or suggestions. Best
would be to file an issue on Github, but email is fine as well.

I hope you find this helpful!
