---
title: "derivmkts: An option pricing package"
output: 
  md_document:
     variant: gfm
---

[![Build Status](https://travis-ci.org/rmcd1024/derivmkts.svg?branch=master)](https://travis-ci.org/rmcd1024/derivmkts)[![](http://www.r-pkg.org/badges/version/derivmkts)](http://www.r-pkg.org/pkg/derivmkts)


```{r, echo=FALSE}
library(knitr)
opts_chunk$set(comment=NA,
               collapse=TRUE)
library(derivmkts)
```


This is a collection of option pricing functions for a course in
financial derivatives. The names of the functions mostly match those
in my book *Derivatives Markets*, which explains the package name.
This should be useful for students comfortable with R. In any case, I
hope it will be helpful for instructors.

There are of course other option pricing packages in R, notably
[RQuantLib](https://cran.r-project.org/package=RQuantLib) and
[fOptions](https://cran.r-project.org/package=RQuantLib). I don't add
significant pricing functionality to those. This package does,
however, have a few aspects that might be unique, which I describe
below: calculation of greeks; display of binomial formula output and
trees; and a quincunx function.

The package includes functions for computing

* Black-Scholes prices and greeks for European options

* Option pricing and plotting using the binomial model

* Barrier options

* Compound options

* Pricing of options with jumps using the Merton model

* Analytical and Monte Carlo pricing of Asian options 


## Calculation of Greeks

The function `greeks()` accepts an option pricing function call as an
argument, and returns a vectorized set of greeks for any pricing
function that uses the input names standard in the package (i.e., the
asset price is `s`, the volatility is `v`, the interest rate is `r`,
the dividend yield is `d`, and the time to maturity is `tt`).

As an example, the following calculation will produce the full
complement of greeks for a call, for each of three strike prices. You
can access the delta values, for example, using `x['Delta', ]`.


```{r}
x1 <- greeks(bscall(s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0))
x1
```

There is a new `complete` option which produces output in a more
convenient format. If `complete=TRUE`, `long=TRUE` will produce
long-form output.


```{r}
x2 <- greeks(bscall(s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0),
             complete=TRUE)
x2

x3 <- greeks(bscall(s=40, k=c(35, 40, 45), v=0.3, r=0.08, tt=0.25, d=0),
             complete=TRUE, long=TRUE)
x3
```

\normalsize


```{r, include=FALSE}
k <- 100; r <- 0.08; v <- 0.30; tt <- 2; d <- 0
S <- seq(.5, 250, by=.5)
y <- list(Call=greeks(bscall(S, k, v, r, tt, d)),
          Put=greeks(bsput(S, k, v, r, tt, d))
          )
par(mfrow=c(4, 4))  ## create a 4x4 plot
par(mar=c(2,2,2,2)) ## shrink some of the margins
for (i in names(y)) {
    for (j in rownames(y[[i]])) {  ## loop over greeks
        plot(S, y[[i]][j, ], main=paste(i, j), ylab=j, type='l')
    }
}
```

The following computes and plots all call and put Greeks for
500 options:


```{r, warnings=FALSE, eval=TRUE}
library(derivmkts)
library(ggplot2)
s <- 100; r <- 0.08; v <- 0.30; tt <- 2; d <- 0
k <- seq(.5, 250, by=.5)
yc <- greeks(bscall(s, k, v, r, tt, d), complete=TRUE, long=TRUE)
yp <- greeks(bsput(s, k, v, r, tt, d), complete=TRUE, long=TRUE)
ggplot(rbind(yc, yp), aes(x=s, y=value, color=funcname)) + geom_line() +
    facet_wrap(~ greek, scales='free_y')
```


## Binomial calculations

### binomopt

By default the binomopt function returns the price of an American
call. In adddition:

* `putopt=TRUE` returns the price of an American
put.

*  `returngreeks=TRUE` returns a subset of the Greeks
along with the binomial parameters.

* `returntrees=TRUE` returns as a list all of the above plus the full
  binomial tree (\$stree), the probability of reaching each node
  (\$probtree), whether or not the option is exercised at each node
  (\$exertree), and the replicating portfolio at each node ($deltatree
  and $bondtree).

Here is an example illustrating everything that the `binomopt`
function can return:
```{r, eval=TRUE}
x <- binomopt(41, 40, .3, .08, 1, 0, 3, putopt=TRUE, american=TRUE,
              returntrees=TRUE)
x
```

### binomplot

This function plots the binomial tree, providing a visual depiction of
the nodes, the probability of reaching each node, and whether exercise
occurs at that node.

```{r}
binomplot(41, 40, .3, .08, 1, 0, 3, putopt=TRUE, american=TRUE)

```

## Galton board or quincunx

The [Galton board](http://mathworld.wolfram.com/GaltonBoard.html) is a
pegboard that illustrates the central limit theorem. Balls drop from
the top and randomly fall right or left, providing a physical
simulation of a binomial distribution. (My physicist brother-in-law
tells me that real-life Galton boards don't typically generate a
normal distribution because, among other things, balls acquire
momentum in the direction of their original travel. The distribution
is thus likely to be fatter-tailed than normal.)

You can see the Galton board in action with `quincunx()`:

```{r, echo=2}
par(mar=c(2, 2, 2, 2))
quincunx(n=11, numballs=250, delay=0, probright=0.5)

```

## Feedback

Please feel free to contact me with bug reports or suggestions. Best
would be to file an issue on Github, but email is fine as well.


