---
title: "An Introduction to R and the derivmkts Package"
author: "Robert McDonald"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  html_document:
    highlight: default
    toc: yes
    toc_float: yes
  html_notebook: default
linkcolor: red
colorlinks: yes
---
	

```{r, echo=FALSE}
library(xtable)
library(knitr)
library(derivmkts)
opts_chunk$set(message=FALSE,
               comment=NA,
               collapse=TRUE,
               prompt=FALSE)
# this is the first document we'll use Jan 24 2017
```


# Overview

## Before we start...

* Open a new "R Notebook" document in RStudio
* Use this to take notes and copy examples
* This will illustrate two things:
    * R, a tool for computations and manipulating and analyzing data
    * Rmarkdown, a tool for creating "statistically literate" documents
* Ask questions at any time if something is not clear!


## Goals

1. Introduction to R: what is it, why learn it?
2. Reproducibility and reporting
3. The derivmkts package


# The steps in data analysis

## The Four Steps of Data Analysis

1. Data acquisition
1. Data "cleaning" (process the data to be researchable)
1. Analysis
1. Reporting (conveying your results)


* R and RStudio provide a reproducible solution for steps 2-4
	* and often step 1 as well...

* Much of your time, and much of the reproducibility challenge, will be in step 2.
* We will focus on basic data manipulation

## The Ford Stock Data Example

* The script you downloaded before class illustrates all four steps in data analysis:
	1. Data acquisition (from Yahoo)
	1. Data formatting (dates,  weekdays, and returns)
	1. Analysis (plots and tables)
	1. Reporting. 


## What is R?

* R is free, open source, statistical software
	* It is a *functional*, *vectorized*, *object-oriented* programming
      language
* R is also part of an ecosystem
    * R the language permits and encourages extensions; it is a development platform
    * The community that writes R packages is large and vibrant
<!--    * Some extensions are commonly used and widely considered "canonical"
        * Packages published by Hadley Wickham and the RStudio team are   examples
        * RCpp permits incorporation of C++ code into R programs
-->
	* R is part of a group of other free and open
      source components
	* This is an important reason there is rapid and enthusiastic
      development surrounding R and RStudio
  * I will say more about this later

## Reproducible and transparent analytics

* It is *essential* to have a data flow that is 
	* robust  <!-- Not too tweaky, not fragile -->
	* reproducible
	* well-documented
	* transparent
* To the extent possible you want it to be scalable
* R facilitates all this


## R vs RStudio

* R is the statistical package
* RStudio is a "front-end" for accessing R
* RStudio incorporates pandoc, for creating different document formats
* If you install RStudio without also installing R, RStudio will be useless
    * However, you can use R without RStudio 


## An example of what you can do with RStudio

* The RStudio folks are rapidly extending the capabilities of R by creating new packages and resources
* RStudio maintains the "shiny" package and also a server to run shiny apps
    * Examples are [here](https://shiny.rstudio.com)
* My babynames shiny page: [https://rmcd1024.shinyapps.io/babynames_shiny](https://rmcd1024.shinyapps.io/babynames_shiny)
* The code for this is on github: [https://github.com/rmcd1024/babynames_shiny](https://github.com/rmcd1024/babynames_shiny)

## Setting expectations

* R and extensions is a huge topic
* The main goal of this workshop is to orient you sufficiently so that you can learn more on your own
* There are wonderful resources on the web,
  including
    * [stackoverflow](http://stackoverflow.com/questions/tagged/r),
    * [R tutorials](http://www.cyclismo.org/tutorial/R/index.html),
    * [Radiant](https://vnijs.shinyapps.io/radiant/) and similar sites
    * [Functions you should know](http://adv-r.had.co.nz/Vocabulary.html)

* Google is your friend


## Making sense of the pieces

* Because of the ecosystem around R 
    * It is powerful
    * It is rapidly-evolving
    * It can be confusing
    
* **Ask** if you don't understand where something fits

## The derivmkts package

* Derivatives analysis in Excel
    * Easy to enter formulas and compute an output
    * What if you want to compute many Greeks?
    * What if you want to distribute your work to others?
    
## Simple option calculations

* Compute one option price

```{r}
library(derivmkts)
s = 100; k=100; v=0.3; r=0.08; tt=1; d=0
bscall(s, k, v, r, tt, d)
```

* Compute greeks

```{r}
greeks(bscall(s, k, v, r, tt, d))
```

## More complex calculations

* Compute many prices (R is vectorized)

```{r}
S = seq(98,102,by=0.5); k=100; v=0.3; r=0.08; tt=1; d=0
bscall(S, k, v, r, tt, d)
```

* Compute many greeks

```{r}
greeks(bscall(S, k, v, r, tt, d))
```

## My favorite example

```{r}
k <- 100; r <- 0.08; v <- 0.30; tt <- 2; d <- 0
S <- seq(.5, 250, by=.5)
Call <- greeks(bscall(S, k, v, r, tt, d))
Put <- greeks(bsput(S, k, v, r, tt, d))
y <- list(Call=Call, Put=Put)
par(mfrow=c(4, 4))  ## create a 4x4 plot
par(mar=c(2,2,2,2))
for (i in names(y)) {
    for (j in rownames(y[[i]])) {  ## loop over greeks
        plot(S, y[[i]][j, ], main=paste(i, j), ylab=j, type='l')
    }
}

```

## Greeks for your own pricing problem

* Suppose you buy a 95 strike call and sell a 105-strike call
* The following will work *provided* that you use "s", "v", "r", "tt", and "d"
to name the respective inputs

```{r}
bullspread = function(s, v, r, tt, d, k1, k2) 
  bscall(s, k1, v, r, tt, d) - bscall(s, k2, v, r, tt, d)
k <- 100; r <- 0.08; v <- 0.30; tt <- 2; d <- 0
S <- seq(.5, 250, by=.5)
k1 = 95; k2 = 105
S=100;
bullspread(S, v, r, tt, d, k1, k2)
greeks(bullspread(S, v, r, tt, d, k1, k2))
```



## Greeks for a bullspread

```{r}
S=seq(0.5, 250, by=0.5);
y = greeks(bullspread(S, v, r, tt, d, k1, k2))
par(mfrow=c(3,3), mar=c(2,2,2,2))
for (j in rownames(y)) {  ## loop over greeks
  plot(S, y[j,], main=j, ylab=j, type='l')
}

```

## Binomial pricing

```{r}
s=100; k=95; r=0.08; tt=1; v=0.30; d=0.08
n = 20;
binomplot(s, k, v, r, tt, d, n=20)
```
```{r}
binomplot(s, k, v, r, tt, d, n=20, plotvalues=TRUE)

```
```{r}
b = binomopt(s, k, v, r, tt, d, n=5, returntrees=TRUE, 
             returnparams=TRUE, returngreeks=TRUE)
b
```



## A final note

* As you start to use R you will likely translate tasks from other environments
(Excel, Stata, SAS, etc.)
* Do *not* try to translate line by line
    * Big temptation, huge mistake
* Instead: 
    * understand the task
    * ask how you would approach it in R
    * look for packages that already do what you want
* *Then* develop your solution in R
